---
title: "Fill Toggle in `ggplot2`"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Data Preparation

First I prepare the data so that relevant variables are factors.
In this example, I will be plotting `hp` on the y-axis and `mpg` on the x-axis.
Color **and** fill will be defined by `cyl` and shape by `gear`.
`vs` is a variable here that will determine whether a data point will be solid or hollow.
I will use `am` as a facet during plotting.

```{r data_prep}
df <- mtcars %>% rownames_to_column("model") %>% 
  mutate(cyl = factor(cyl),
         vs  = ifelse(vs, "Straight", "V-shaped"),
         gear = factor(gear),
         am = factor(am)) %>%
  as_tibble()

glimpse(df)
summary(df)
```

## Plotting

Below I initialize the plot and assign it to `pl`.
I begin by setting the `aes` mappings for `x`, `y`, `color`, and `shape`.
Next I add a `geom_point` that has extra mappings of `fill` and `alpha`.
Note that `fill` and `color` for this geom layer are mapped to the same column, `cyl`.
`alpha`, which controls the transparency, is mapped to a binary variable
(only has 0 and 1 values) and is converted to a character.
This means that if `vs == 1` then this point will be fully visible and if `vs == 0`
it will be invisible.
Finally, a `geom_point` layer is added to the plot that can only have unfilled points.


```{r initial_plot}
pl <- ggplot(df,
       aes(
         x = mpg,
         y = hp,
         color = cyl,
         shape = gear
       )) +
  geom_point(aes(
    fill = cyl,
    alpha = vs
  )) +
  geom_point()

```

We next need to specify a few things for this all to work.
First in `scale_alpha_manual`, I specify the values using a named vector.
In `scale_shape_manual`, I specify the shapes that this is allowed to be
(only those that can have both color and fill) and rename the legend.
In the `guides` layer, I specify that the elements in the `shape` legend should
be filled in and that `alpha` legend keys should be fully visible,
simulating the empty shapes by setting fill to `c("white", "black")`.
Do note that the order here is important, it must follow the order of `values`
in `scale_alpha_manual` or the legend will confuse the reader.

```{r scales_guides}
pl +
  scale_alpha_manual(values = c("V-shaped" = 0, "Straight" = 1),
                     breaks = c("V-shaped", "Straight"),
                     name = "Engine") +
  scale_shape_manual(values = 21:25,
                     name = "Number of forward gears") +
  guides(shape = guide_legend(override.aes = list(fill = "black")),
         alpha = guide_legend(override.aes = list(alpha = c(1,1),
                                                  fill = c("white", "black"),
                                                  shape = 21)) ) +
  facet_wrap(~am) +
  theme_bw()
```

