---
title: "Plotting Model Fits"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plotting modelling functions with observed data

This will have 2 models analyzed 2 different ways
Model1 = How much volume has a cylinder lost at a certain time?
Model1 parameters = radius, initial height, height rate of change
Model2 = How tall is liquid filling a semi-spherical after a certain time?
Model2 parameters = max radius, volume rate of change

Model1 is given by:

$$
f(t) = \pi r^2h_0 - \pi r^2\frac{dh}{dt}\cdot t 
$$

Model2 is given by:

$$
g(t) = \frac{dV}{dt}\cdot\frac{t}{2\pi r^2}
$$

First I will make a simple data.frame with simple parameter entries for each fit.
The way they are fit are arbitrary, I just want to see how I can combine it later with
the observed data into one plot.
There are three situations: one where only Model1 is fit, one where both Models are fit,
and one where only Model2 is fit.
So there are 2 x 3 = 6 rows in the fits.

For the observed data, lets say that there is four nonstandard time-points per situation,
for each type of measurements.
There are four measurement events by four time-points... meaning 16 different observations.

The values are volume or rate of change for radius for Model1 and Model2, respectively.

```{r}
# Load Packages ----
{
library(tidyverse)
library(gridExtra)
set.seed(1010)
}

# Initial height for Model1 is constant at 20
height_init <- 20

# Recorded fit parameters -----
fit_df <- data.frame(
  Situation = rep(LETTERS[1:3], each = 2),
  Fit_type = rep(c("Arb1", "Arb2"), 3),
  radius.Model1 = c(rnorm(2, 5), rnorm(2, 4.5), NA, NA),
  height_rc.Model1 = c(rnorm(2, 0.25, 0.1), rnorm(2, 0.5, 0.1), NA, NA),
  radius.Model2 = c(NA, NA, rnorm(2, 1, 0.1), rnorm(2, 1.4, 0.2)),
  volume_rc.Model2 = c(NA, NA, rnorm(2, 1, 0.02), rnorm(2, 1.3, 0.02))
)


obs_df <- data.frame(
  Situation = c(rep("A", 4), rep("B", 8), rep("C", 4)),
  Model = c(rep("Model1", 8), rep("Model2", 8)),
  Measurement = c(60, 50, 40, 30,
                  48, 40, 35, 25,
                  0, 0.1, 0.2, 0.4,
                  0, 0.2, 0.5, 0.6),
  Timepoint = c(rep(c(0, 2, 4, 8), 2), rep(c(0, 1, 2, 3), 2))
)

```

Now we have our two data sets.
Next we try to split the Model and Estimates in `fit_df`.

```{r}
fit_df <- fit_df %>%
  pivot_longer(cols = contains(".Model"),
               names_to = "EstimateName.FitModel",
               values_to = "Estimate") %>%
  separate_wider_delim(cols = EstimateName.FitModel,
                       delim = ".",
                       names = c("EstimateName", "FitModel"))

```

At this point `fit_df` is ready to be merged with the observation data.
It must then be filtered by where Model == FitModel and nested by the Observation data
and the FitEstimates.
Finally, the estimates must be pivoted so that each parameter has it's own column.

```{r}
combo_df <- fit_df %>% left_join(obs_df, by = "Situation",
                                 relationship = "many-to-many") %>%
  filter(Model == FitModel) %>%
  nest(observations = c(Measurement, Timepoint),
       fits = c(EstimateName, Estimate)) %>%
  mutate(fits = map(fits, distinct), # Values were doubled during the join
         fits = map(fits, deframe), # deframe converts two-column data.frames into named list
         fits = map(fits, unlist),
         fits = map(fits, as.list),
         fits = map(fits, as_tibble),
         param_model = paste(Fit_type,
                             FitModel)) %>%
  group_by(Situation, Fit_type, Model) # These columns uniquely identify the columns
# The above is incredibly important!!!

# Example
combo_df$fits[[1]]
```

Once our data is in order, I will write the functions that define each of our models.
In the chunk below they are displayed as plots using the `stat_function` function.

```{r}

# Model1
ggplot() +
  stat_function(
    fun = function(t, radius, height_roc) {
      pi*(radius^2)*(height_init - (height_roc*t))
      }, args = list(radius = 1, height_roc = 0.5)
  ) +
  xlim(0, 10)

# Model2
ggplot() +
  stat_function(
    fun = function(t, radius, volume_roc) {
      volume_roc * t / (2 * pi * (radius^2))
      }, args = list(radius = 1, volume_roc = 1.2)
  ) +
  xlim(0, 10)
```

If we generalize these to just a formula.

```{r}
m1_fun <- function(t, radius, height_roc) {
      pi*(radius^2)*(height_init - (height_roc*t))
}

m2_fun <- function(t, radius, volume_roc) {
      volume_roc * t / (2 * pi * (radius^2))
}

```


Finally, it is time to combine these two using `map2` from the `purrr` package.

```{r}

combined_plots <- combo_df %>%
  mutate(combo_plot = map2(.x = observations,
                           .y = fits,
                           .f = (\(.x, .y) {
                             ggplot(data = .) +
                               geom_point(data = .x,
                                          aes(x = Timepoint, y = Measurement)) +
                               {
                                 if (Model == "Model1") {
                                   stat_function(fun = function(tp, radius, height_roc) {
                                     pi * (radius^2) * (20 - (height_roc * tp))
                                   },
                                   args = list(
                                     radius = .y[["radius"]],
                                     height_roc = .y[["height_rc"]])
                                   )
                                 } else {
                                   stat_function(fun = function(tp, radius, volume_roc) {
                                     volume_roc * tp / (2 * pi * (radius^2))
                                   },
                                   args = list(
                                     radius = .y[["radius"]],
                                     volume_roc = .y[["volume_rc"]])
                                   )
                                 }
                               } + 
                               labs(title = paste(Situation, Model)) +
                               xlim(0, 10) +
                               theme_bw() +
                               theme(aspect.ratio = 1)
                           }
                           ), .progress = TRUE
  ))

# Issue with if (model == "Model1") conditon has length > 1
# Does ifelse solve this? No, it needed to be grouped by unique identifying columns

# However, I think now that the different arguments for Model1 and Model2 are causing
# a recycling error! :(

# Trying to filter out just Model1 to see if that works (no conditional statement)
combined_plots <- combo_df %>%
  filter(Model == "Model1") %>%
  select(-Fit_type, -param_model) %>% distinct() %>%
  mutate(combo_plot = map2(.x = observations,
                           .y = fits,
                           .f = (\(.x, .y) {
                             ggplot(data = .) +
                               geom_point(data = .x,
                                          aes(x = Timepoint,
                                              y = Measurement)
                               ) +
                               stat_function(
                                 fun = function(x, radius, height_roc) {
                                   pi * (radius^2) * (20 - (height_roc * x))
                                 },
                                 args = list(
                                   radius = .y[["radius"]],
                                   height_roc = .y[["height_roc"]])
                               ) + 
                               labs(title = paste(Situation, Model)) +
                               xlim(0, 10) +
                               theme_bw()
                           }
                           )
  ))
# Still getting the recycling error
```

Next I need to combine all the plots saved in `combo_plot` in a list.
For this I will use `summarize`.


```{r}
grid_combo <- combined_plots %>%
  group_by(Situation) %>%
  summarize(grid_plots = list(combo_plot))

for (i in 1:nrow(grid_combo)) {
  grid.arrange(grobs = grid_combo$grid_plots[[i]],
               nrow = 1,
               top = paste("This is ",
                           grid_combo$Situation[[i]]))
}

```

